\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

% ── Packages ──────────────────────────────────────────────────────────
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{balance}  % balance columns on last page

% ── Macros ────────────────────────────────────────────────────────────
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

% Graphics path
\graphicspath{{./figures/}}

% ── Document ──────────────────────────────────────────────────────────
\begin{document}

\title{Prompt Injection and Tool-use Attacks in LLM Applications: A Survey}

% ── Authors (5 members) ──────────────────────────────────────────────
\author{
\IEEEauthorblockN{Yu Peihan}
\IEEEauthorblockA{Dept. of Computer Science \\
City University of Hong Kong (Dongguan) \\
ID: 72510479 \\
YuFishYPH@gmail.com}
\and
\IEEEauthorblockN{Hu Xuelin}
\IEEEauthorblockA{Dept. of Computer Science \\
City University of Hong Kong (Dongguan) \\
ID: 72510295 \\
72510295@cityu-dg.edu.cn}
\and
\IEEEauthorblockN{Weng Ziyan}
\IEEEauthorblockA{Dept. of Computer Science \\
City University of Hong Kong (Dongguan) \\
ID: 72510283 \\
72510283@cityu-dg.edu.cn}
\and
\IEEEauthorblockN{Deng Leying}
\IEEEauthorblockA{Dept. of Computer Science \\
City University of Hong Kong (Dongguan) \\
ID: 72512227 \\
72512227@cityu-dg.edu.cn}
\and
\IEEEauthorblockN{Zhao Haichao}
\IEEEauthorblockA{Dept. of Computer Science \\
City University of Hong Kong (Dongguan) \\
ID: 72515790 \\
72515790@cityu-dg.edu.cn}
}

\maketitle

% ── Abstract ──────────────────────────────────────────────────────────
\begin{abstract}
Large Language Models (LLMs) are increasingly deployed in applications that
process untrusted inputs and invoke external tools, creating new attack surfaces.
This survey systematically examines prompt injection attacks---both direct and
indirect---as well as tool-use exploitation in LLM-integrated systems. We
propose a unified taxonomy covering three attack families (direct prompt
injection, indirect prompt injection, and tool-use/agent exploitation) and four
defense categories (input sanitization, instruction hierarchy, output
monitoring, and architectural isolation). Through critical analysis of recent
literature, we compare threat model assumptions, evaluation methodologies, and
the practical deployability of proposed defenses. Our findings reveal that
current defenses remain largely reactive and brittle, and we identify key open
problems including the absence of standardized benchmarks and the fundamental
tension between model capability and security.
\end{abstract}

% ── Section files ─────────────────────────────────────────────────────
\input{sections/introduction}
\input{sections/background}
\input{sections/analysis_attacks}
\input{sections/analysis_defenses}
\input{sections/analysis_comparison}
\input{sections/discussion}
\input{sections/conclusion}
\input{sections/ai_statement}

% ── References ────────────────────────────────────────────────────────
\balance
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}
