% ── AI Use Statement ─────────────────────────────────────────────────
\section*{AI Use Statement}

\textbf{Tools Used.}
The following AI tools were used during the preparation of this report:
\begin{itemize}
  \item \textbf{ChatGPT (GPT-4o, OpenAI)} --- used for literature search, language polishing, and reference verification.
  \item \textbf{Claude (Anthropic)} --- used for literature search and citation cross-checking.
\end{itemize}

\textbf{How AI Was Used.}
AI tools served a \emph{supporting} role in the following specific tasks:

\begin{enumerate}
  \item \textbf{Literature Search (AI-assisted):} AI tools helped generate search queries for Google Scholar and Semantic Scholar, and produced initial one-paragraph summaries of candidate papers to accelerate the screening process.
  \item \textbf{Report Outline and Core Arguments (Human-authored):} The overall report structure, analytical framework, attack--defense taxonomy, critical arguments, and cross-paper synthesis were designed and written entirely by the team members. AI tools were \emph{not} used to determine the report's logical flow or argumentative direction.
  \item \textbf{Language Polishing (AI-assisted):} After the team completed each section draft in our own words, AI tools were used to improve grammar, phrasing, and readability. All polished text was reviewed and approved by the original author of each section.
  \item \textbf{Reference Verification (AI-assisted):} AI tools were used to cross-check BibTeX entries against original papers (author names, titles, publication venues, years) and to verify that in-text citation content accurately reflected the cited source.
\end{enumerate}

\textbf{Representative Prompts.}
\begin{enumerate}
  \item ``Search for recent papers on indirect prompt injection attacks targeting LLM-integrated applications published after 2022.''
  \item ``Polish the following paragraph for academic English while preserving the original meaning: [paragraph pasted].''
  \item ``Check whether the following BibTeX entry matches the information on the paper's first page: [entry pasted].''
  \item ``Verify that the claim `InjecAgent reports a 23.9\% unsafe-action rate' is accurately attributed to Zhan et al.\ (2024).''
\end{enumerate}

\textbf{Verification Process.}
Every AI-generated summary was cross-checked against the original paper. No AI-produced text was included in the final report without rewriting by a team member. Factual claims (e.g., specific numbers such as attack success rates) were verified by locating the corresponding table or figure in the cited source. Any AI output that could not be verified was discarded.

\textbf{Academic Integrity.}
We confirm that this report represents our own analysis and synthesis of the surveyed literature. The core intellectual contributions---the analytical framework, critical evaluations, comparative arguments, and original insights---are entirely the work of the five team members. AI tools were used as assistants for searching, polishing, and checking, not as authors.
